---
title: "Assignment"
output: html_document
---

Part 1 

```{r}
library('rpart')
library('DAAG')
library('dismo')
library('randomForest')
library("e1071")

sonar.df = read.table("sonar.all-data.txt", header = FALSE, sep = ",", na.strings = "NA" , dec = ".", strip.white = FALSE)
```

Here, I used the rpart method instead of a C4.5 classifier due to problems with installation

```{r}
tree <- rpart(V61~., data= sonar.df, na.action = na.rpart, x = TRUE, y = TRUE)
predictions <- predict(tree, sonar.df[, 1:60], type="class")
summary(tree)

table(predictions, sonar.df$V61)
confusionMatrix <- table(predictions, sonar.df$V61)

```

My function to comute error, accuracy and precision

```{r}
computeErrorMetrics <- function(confusionMatrix) {
p = confusionMatrix[1,1]/ (confusionMatrix[1,1] + confusionMatrix[2,1])

cat("precision is", p)
r = confusionMatrix[1,1]/ (confusionMatrix[1,1] + confusionMatrix[1,2])

cat("Recall is", r)

a = (confusionMatrix[1,1] + confusionMatrix[2,2]  )/ (confusionMatrix[1,1] + confusionMatrix[1,2] + confusionMatrix[2,1] + confusionMatrix[2,2] )

cat("Accuracy is", a)

error = 1 - a

cat("Error is", a)
fscore = (2*p*r)/ (p + r)
cat("fscore is", fscore)

}
```
Obviously, the accuracy is high due to overfitting


```{r}

computeErrorMetrics(confusionMatrix)

```



Kfold, Rpart



```{r}
folds <- kfold(sonar.df, k=10 )

prediction <- data.frame()
testset <- data.frame()

for(i in 1:10){
  test <- sonar.df[folds==i,]
  train <- sonar.df[folds!=i,]
  model <- rpart(V61~., data= train, na.action = na.rpart, x = TRUE, y = TRUE)
  testset<- rbind(testset, as.data.frame(test))
  predictedFold <- data.frame(predict(model, test[, 1:60], type="class"))
  prediction <- rbind(prediction, predictedFold)
}

cm <- table(as.factor(prediction$predict.model..test...1.60...type....class..), testset$V61)

computeErrorMetrics(cm)
```



Random Forest using K-fold

```{r}
prediction.rf <- data.frame()
testset.rf <- data.frame()

for(i in 1:10){
  test.rf <- sonar.df[folds==i,]
  train.rf <- sonar.df[folds!=i,]
  model.rf <- randomForest(V61~., data= train.rf)
  testset.rf<- rbind(testset.rf, as.data.frame(test.rf))
  predictedFold.rf <- data.frame(predict(model.rf, test.rf[, 1:60], type="class"))
  prediction.rf <- rbind(prediction.rf, predictedFold.rf)
}

cm.rfk <- table(prediction.rf$predict.model.rf..test.rf...1.60...type....class.., testset.rf$V61)

computeErrorMetrics(cm.rfk)

```



```{r}
prediction.rf <- data.frame()
testset.rf <- data.frame()

for(i in 1:10){
  test.rf <- sonar.df[folds==i,]
  train.rf <- sonar.df[folds!=i,]
  model.rf <- randomForest(V61~., data= train.rf)
  testset.rf<- rbind(testset.rf, as.data.frame(test.rf))
  predictedFold.rf <- data.frame(predict(model.rf, test.rf[, 1:60], type="class"))
  prediction.rf <- rbind(prediction.rf, predictedFold.rf)
}

cm.rfk <- table(prediction.rf$predict.model.rf..test.rf...1.60...type....class.., testset.rf$V61)

computeErrorMetrics(cm.rfk)

```

SVM
```{r}
prediction.svm <- data.frame()
testset.svm <- data.frame()

for(i in 1:10){
  test.svm <- sonar.df[folds==i,]
  train.svm <- sonar.df[folds!=i,]
  model.svm <- svm(V61~., data= train.svm)
  testset.svm<- rbind(testset.svm, as.data.frame(test.svm))
  predictedFold.svm <- data.frame(predict(model.svm, test.svm[, 1:60], type="class"))
  prediction.svm <- rbind(prediction.svm, predictedFold.svm)
}

cm.svmk <- table(prediction.svm$predict.model.svm..test.svm...1.60...type....class.., testset.svm$V61)

computeErrorMetrics(cm.svmk)
```
naiveBayes

```{r}
prediction.nb <- data.frame()
testset.nb <- data.frame()

for(i in 1:10){
  test.nb <- sonar.df[folds==i,]
  train.nb <- sonar.df[folds!=i,]
  model.nb <- naiveBayes(V61~., data= train.nb)
  testset.nb<- rbind(testset.nb, as.data.frame(test.nb))
  predictedFold.nb <- data.frame(predict(model.nb, test.nb[, 1:60], type="class"))
  prediction.nb <- rbind(prediction.nb, predictedFold.nb)
}

cm.nbk <- table(prediction.nb$predict.model.nb..test.nb...1.60...type....class.., testset.nb$V61)

computeErrorMetrics(cm.nbk)
```

for hepatitis, reading files

```{r}
hepatitis.df <- read.table("hepatitis.data.txt", header = FALSE, sep = ",", na.strings = "?" , dec = ".", strip.white = FALSE)
```

Perform rpart on Hepatitis

```{r}
folds <- kfold(hepatitis.df, k=10 )

prediction <- data.frame()
testset <- data.frame()

for(i in 1:10){
  test <- hepatitis.df[folds==i,]
  train <- hepatitis.df[folds!=i,]
  model <- rpart(V20~., data= train, na.action = na.rpart, x = TRUE, y = TRUE)
  testset<- rbind(testset, as.data.frame(test))
  predictedFold <- data.frame(predict(model, test[, 1:19]))
  prediction <- rbind(prediction, predictedFold)
}

cm <- table(as.factor(prediction$predict.model..test...1.19..), testset$V20)

computeErrorMetrics(cm)

```

Random Forest Hepatitis
Random Forest has a problem with missing values, so it has to be handled, before applying the method
```{r}
prediction.rf <- data.frame()
testset.rf <- data.frame()
for(i in 1:10){
  test.rf <- hepatitis.df[folds==i,]
  train.rf <- hepatitis.df[folds!=i,]
  model.rf <- randomForest(as.factor(V20)~., data= train.rf, na.action = na.omit )
  testset.rf<- rbind(testset.rf, as.data.frame(test.rf))
  predictedFold.rf <- data.frame(predict(model.rf, test.rf[, 1:19]))
  prediction.rf <- rbind(prediction.rf, predictedFold.rf)
}
cm.rfk <- table(prediction.rf$predict.model.rf..test.rf...1.19.., testset.rf$V20)
computeErrorMetrics(cm.rfk)

```




Naive bayes Hepatitis


```{r}
prediction.nb <- data.frame()
testset.nb <- data.frame()

for(i in 1:10){
  test.nb <- hepatitis.df[folds==i,]
  train.nb <- hepatitis.df[folds!=i,]
  model.nb <- naiveBayes(as.factor(V20)~., data= train.nb)
  testset.nb<- rbind(testset.nb, as.data.frame(test.nb))
  predictedFold.nb <- data.frame(predict(model.nb, test.nb[, 1:19]))
  prediction.nb <- rbind(prediction.nb, predictedFold.nb)
}

cm.nbk <- table(as.factor(prediction.nb$predict.model.nb..test.nb...1.19..), testset.nb$V20)

computeErrorMetrics(cm.nbk)

```


SVM hepatitis

```{r}
hepatitis.df <- read.table("hepatitis.data.txt", header = FALSE, sep = ",", na.strings = "NA" , dec = ".", strip.white = FALSE)
testset.sv <- data.frame()

prediction.sv <- data.frame()
for(i in 1:10){
  test.sv <- hepatitis.df[folds==i,]
  train.sv <- hepatitis.df[folds!=i,]
  model.sv <- svm(as.factor(V20)~., data= train.sv)
  testset.sv<- rbind(testset.sv, as.data.frame(test.sv))
  predictedFold.sv <- data.frame(predict(model.sv, test.sv[, 1:19]))
  prediction.sv <- rbind(prediction.sv, predictedFold.sv)
}

cm.svmk <- table(prediction.sv$predict.model.sv..test.sv...1.19.., testset.sv$V20)

computeErrorMetrics(cm.svmk)



```


Pima data
```{r}
pima.df = read.table("pima-indians-diabetes.data.txt", header = FALSE, sep = ",", na.strings = "NA" , dec = ".", strip.white = FALSE)

```

Pima Rpart

```{r}
folds <- kfold(pima.df, k=10 )

prediction <- data.frame()
testset <- data.frame()

for(i in 1:10){
  test <- pima.df[folds==i,]
  train <- pima.df[folds!=i,]
  model <- rpart(V9~., data= train, na.action = na.rpart, x = TRUE, y = TRUE)
  testset<- rbind(testset, as.data.frame(test))
  predictedFold <- data.frame(predict(model, test[, 1:8]))
  prediction <- rbind(prediction, predictedFold)
}

cm <- table(as.factor(prediction$predict.model..test...1.8..), testset$V9)

computeErrorMetrics(cm)

```

Pima random forest

```{r}

prediction.rf <- data.frame()
testset.rf <- data.frame()

for(i in 1:10){
  test.rf <- pima.df[folds==i,]
  train.rf <- pima.df[folds!=i,]
  model.rf <- randomForest(as.factor(V9)~., data= train.rf)
  testset.rf<- rbind(testset.rf, as.data.frame(test.rf))
  predictedFold.rf <- data.frame(predict(model.rf, test.rf[, 1:8]))
  prediction.rf <- rbind(prediction.rf, predictedFold.rf)
}
cm.rfk <- table(prediction.rf$predict.model.rf..test.rf...1.8.., testset.rf$V9)

computeErrorMetrics(cm.rfk)

```

Pima SVM

```{r}

prediction.svm <- data.frame()
testset.svm <- data.frame()

for(i in 1:10){
  test.svm <- pima.df[folds==i,]
  train.svm <- pima.df[folds!=i,]
  model.svm <- svm(V9~., data= train.svm)
  testset.svm<- rbind(testset.svm, as.data.frame(test.svm))
  predictedFold.svm <- data.frame(predict(model.svm, test.svm[, 1:8]))
  prediction.svm <- rbind(prediction.svm, predictedFold.svm)
}

cm.svmk <- table(prediction.svm$predict.model.svm..test.svm...1.8.., testset.svm$V9)

computeErrorMetrics(cm.svmk)

```

Naive bayes pima


```{r}
prediction.nb <- data.frame()
testset.nb <- data.frame()

for(i in 1:10){
  test.nb <- pima.df[folds==i,]
  train.nb <- pima.df[folds!=i,]
  model.nb <- naiveBayes(as.factor(V9)~., data= train.nb)
  testset.nb<- rbind(testset.nb, as.data.frame(test.nb))
  predictedFold.nb <- data.frame(predict(model.nb, test.nb[, 1:8]))
  prediction.nb <- rbind(prediction.nb, predictedFold.nb)
}

cm.nbk <- table(prediction.nb$predict.model.nb..test.nb...1.8.., testset.nb$V9)

computeErrorMetrics(cm.nbk)

```

Spect data
```{r}
spect.train.df = read.table("SPECT.train.txt", header = FALSE, sep = ",", na.strings = "NA" , dec = ".", strip.white = FALSE)
spect.test.df = read.table("SPECT.test.txt", header = FALSE, sep = ",", na.strings = "NA" , dec = ".", strip.white = FALSE)

spect.df = rbind(spect.train.df, spect.test.df)
```


Spect Rpart

```{r}
folds <- kfold(spect.df, k=10 )

prediction <- data.frame()
testset <- data.frame()

for(i in 1:10){
  test <- spect.df[folds==i,]
  train <- spect.df[folds!=i,]
  model <- rpart(V23~., data= train, na.action = na.rpart, x = TRUE, y = TRUE)
  testset<- rbind(testset, as.data.frame(test))
  predictedFold <- data.frame(predict(model, test[, 1:22]))
  prediction <- rbind(prediction, predictedFold)
}

cm <- table(as.factor(prediction$predict.model..test...1.22..), testset$V23)

computeErrorMetrics(cm)

```

spect random forest

```{r}

prediction.rf <- data.frame()
testset.rf <- data.frame()

for(i in 1:10){
  test.rf <- spect.df[folds==i,]
  train.rf <- spect.df[folds!=i,]
  model.rf <- randomForest(as.factor(V23)~., data= train.rf)
  testset.rf<- rbind(testset.rf, as.data.frame(test.rf))
  predictedFold.rf <- data.frame(predict(model.rf, test.rf[, 1:22]))
  prediction.rf <- rbind(prediction.rf, predictedFold.rf)
}
cm.rfk <- table(prediction.rf$predict.model.rf..test.rf...1.22.., testset.rf$V23)

computeErrorMetrics(cm.rfk)

```

spect SVM

```{r}

prediction.svm <- data.frame()
testset.svm <- data.frame()

for(i in 1:10){
  test.svm <- spect.df[folds==i,]
  train.svm <- spect.df[folds!=i,]
  model.svm <- svm(V23~., data= train.svm)
  testset.svm<- rbind(testset.svm, as.data.frame(test.svm))
  predictedFold.svm <- data.frame(predict(model.svm, test.svm[, 1:22]))
  prediction.svm <- rbind(prediction.svm, predictedFold.svm)
}

cm.svmk <- table(prediction.svm$predict.model.svm..test.svm...1.22.., testset.svm$V23)

computeErrorMetrics(cm.svmk)

```

Naive bayes spect


```{r}
prediction.nb <- data.frame()
testset.nb <- data.frame()

for(i in 1:10){
  test.nb <- spect.df[folds==i,]
  train.nb <- spect.df[folds!=i,]
  model.nb <- naiveBayes(as.factor(V23)~., data= train.nb)
  testset.nb<- rbind(testset.nb, as.data.frame(test.nb))
  predictedFold.nb <- data.frame(predict(model.nb, test.nb[, 1:22]))
  prediction.nb <- rbind(prediction.nb, predictedFold.nb)
}

cm.nbk <- table(prediction.nb$predict.model.nb..test.nb...1.22.., testset.nb$V23)

computeErrorMetrics(cm.nbk)
```
